Flowchart of Current Algorithm

### 2009-04-27 docs/flowchart/flowchart.tex
somewhat major changes to integrate the volume discrepancy correction term in a more seamless manner

\section{Changes Between Paper 3 and Paper 4}
\begin{enumerate}
\item {\tt fill\_3d} and {\tt average} algorithms.
\item We update temperature after each {\bf React State} as well as the advection step.
\item We have created a {\tt burning\_cutoff\_density}, where the burning does
not happen below this density.  It is presently set to {\tt base\_cutoff\_density}.
\item Use corner coupling in advection.
\item We use {\tt use\_tfromp} to update temperature using $T=T(\rho,X_k,p_0)$ rather 
than $T=T(\rho,h,X_k)$.
\end{enumerate}
Changes which modify enthalpy update.  These are separate because in this paper,
enthalpy is only used as a diagnostic since we use {\tt use\_tfromp}.  If we don't
mention these items in paper IV, they should me noted in paper 4.5.
\begin{enumerate}
\item We have changed the discretization of $\Ub\cdot\nabla p_0$ in the enthalpy
update to $\nabla\cdot(\Ub p_0) - p_0\nabla\cdot\Ub$.
\item In paper III we discretized the enthalpy evolution equation in
full, non-perturbation form, $\rho h$.  Since then we have discovered that 
discretizing the enthalpy evolution in perturbational form, $(\rho h)'$,
leads to better numerical properties.  We use {\tt enthalpy\_pred\_type = 1}.
This is more like paper II.
\item We have turned off the evolution of $h$ above the atmosphere and instead
compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff = TRUE}.
\end{enumerate}
\section{Changes Between Papers 4 and Papers 4.5}
\begin{enumerate}
\item We split the update of the density and enthalpy in steps 4 and 8.
Now, density is updated first, allowing us to compute time-centered values of 
$\psi$ that are then used in the enthalpy update.
\item We split {\bf Advect Base} into separate {\bf Advect Base Density},
{\bf Advect Base Pressure}, and {\bf Advect Base Enthalpy}.  The calling sequence
is in the main flow chart.
\item We replaced {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
plane-parallel and $\rho_0 = \rho_0 + {\rm{\bf Avg}}(\rho - \rho_0)$ in spherical,
eliminating the need for $\nabla\cdot\etarho$.  {\bf Correct Base} is now used
analogously for setting $(\rho h)_0$ after {\bf React State}.
\item After {\bf React State}, we set $(\rho h)_0 = {\rm{\bf Avg}}(\rho h)$ in 
plane-parallel and $(\rho h)_0 = (\rho h)_0 + {\rm{\bf Avg}}[(\rho h) - (\rho h)_0]$ 
in spherical.  This means that {\bf React Base} no longer exists, and the intermediate 
states, $(\rho h)_0^{(1)}$ and $(\rho h)_0^{(2)}$, no longer have to exist.  
{\bf Advect Base Enthalpy} will advance $(\rho h)_0^n \rightarrow (\rho h)_0^{n+1}$ 
directly.  This also means we don't have to deal with $\eta_{\rho h}$ or a 
{\bf Diffuse Base}.
\item In {\bf Step 4} and {\bf Step 8}, we moved {\bf Correct Base} so it
occurs after the full state density update and  before {\bf Advect Base Pressure}.  
Since the full enthalpy advance ``feels'' the effects of the time-centered $\etarho$, 
we should rely on a $\rho_0^{n+1}$ that also feels the effects.
\item {\bf AMR!!!}
\end{enumerate}
Changes specific to spherical:
\begin{enumerate}
\item Discuss {\bf Advect Base} routines and {\bf Correct Base} routine.
\item $\etarho$ and $\psi$ equations.
\end{enumerate}
\section{Changes Between Papers 4.5 and Paper 5}
\begin{enumerate}
\item Use PPM limiters in advection.
\end{enumerate}
\section{Changes Between Paper 5 and XRB Paper}
\begin{enumerate}
\item We have added thermal diffusion, controlled by {\tt use\_thermal\_diffusion},
{\tt temp\_diffusion\_formulation}, and {\tt thermal\_diffusion\_type}.
\item We added the volume-discrepancy term to the velocity constraint equation,
controlled by the input parameter, {\tt dpdt\_factor}.
\end{enumerate}

@@

We integrate
\begin{equation}
\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \psi
\enskip \label{eq:divw0} ,
\end{equation}
with the following discretization:
\begin{equation}
\frac{\partial w_0^{\outp}}{\partial r} =  \Sbar^{\inp}
- \frac{1}{\gammabar^{\inp} p_0^{\inp}}\psi^{\inp}
\enskip ,
\end{equation}
and $w_0=0$ at $r=0$.
% VOLUME DISCREPANCY STUFF
%\begin{equation}
%\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S^{\nph,\star}}
%- \frac{1}{\gammabar^{n} p_0^{n}}
%\left\{ \psi^{\nmh} + 
%f \left[\frac{p_0^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n} \right]\right\}
%\enskip .
%\end{equation}

@@

If this is the first step of the calculation ($n=0$), we set
\begin{equation}
S^{\myhalf,\star} = \frac{S^0 + S^1}{2} \enskip ,
\end{equation}
where $S^1$ is found through the iterative process that initializes the calculation.
Otherwise, following the method used in our small-scale low Mach number algorithm
(BDRWZ), we extrapolate to the half-time using the source
terms at the previous and current time levels
\begin{equation}
S^{\nph,\star} = S^n + \frac{\Delta t^n}{2} \frac{S^n - S^{n-1}}{\Delta t^{n-1}} \enskip .
\end{equation}
\item Compute
\begin{equation}
\overline{S^{\nph,\star}} = {\mathrm{\bf Avg}} \left(S^{\nph,\star}\right) \enskip .
\end{equation}
\item Compute $w_0^{n+\myhalf,\star}$:
\begin{description}
\item For plane-parallel,\\
{\bf Compute $w_0$ Planar}$[\overline{S^{\nph,\star}},\gammabar^n,p_0^n,\psi^{n-\myhalf}] 
\rightarrow [w_0^{n+\myhalf,\star}]$.
\item For spherical,\\
{\bf Compute $w_0$ Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,\rhozero^n,p_0^n,
\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.
\end{description}

@@

The procedure to construct $\uadvone$ is described in detail in Appendix B of paper III.
We note that  $\uadvone$ satisfies the discrete versions of paper III equation (21),
\begin{equation}
\int_{\Omega_H} \Ubt \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
\end{equation}
and paper III equation (39),
\begin{equation}
\nablab \cdotb (\beta_0 \Ubt )  = \beta_0 \left(S - \Sbar \right )\enskip .
\label{eq:tildeconstraint}
\end{equation}
The discretization with the volume discrepancy correction is
\begin{equation}
\nablab \cdotb \left(\beta_0^n \uadvone\right) = 
\beta_0^n \left\{ \left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right)
+ \frac{f}{\gammabar^n p_0^n}
\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right]\right\} \enskip .
\end{equation}

@@

\item Define:
\begin{equation}
\gammabar^{\nph,\star} = \frac{\gammabar^{n} + \gammabar^{n+1,\star}}{2}\enskip , ~~~
\rhozero^{\nph,\star} = \frac{\rhozero^{n} + \rhozero^{n+1,\star}}{2}\enskip , ~~~
p_0^{\nph,\star} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}\enskip .
\end{equation}
\item Compute $w_0^{n+\myhalf}$:
\begin{description}
\item For plane-parallel,\\
{\bf Compute $w_0$ Planar}$[\overline{S^{\nph}},
\gammabar^{n+\myhalf,\star},p_0^{n+\myhalf,\star},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf,\star}]$.
\item For spherical,\\
{\bf Compute $w_0$ Spherical}$[\overline{S^{\nph}},\gammabar^{n+\myhalf,\star},
\rhozero^{n+\myhalf,\star},p_0^{n+\myhalf,\star},\etarho^{n+\myhalf,\star}] \rightarrow [w_0^{n+\myhalf}]$.
\end{description}
% VOLUME DISCREPANCY STUFF
%\begin{equation}
%\overline{p(\rho,h,X_k)^{\nph,\star}} = 
%\frac{\overline{p(\rho,h,X_k)^n} + \overline{p(\rho,h,X_k)^{n+1,\star}}}{2}\enskip .
%\end{equation}
%\begin{equation}
%\frac{\partial w_0^{\nph}}{\partial r} = 
%\overline{S^{\nph}} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
%\left\{ \psi^{\nph,\star} + f \left[\frac{p_0^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right]\right\}
%\enskip ,

\item Using equation (\ref{eq:pizero}), define the scaled pressure gradient using the
following discretization:
\begin{equation}
-\left ( \frac{1}{\rho_0^n} \frac{\partial \pizerotwo}{\partial r} \right ) = 
\frac{w_0^{\nph} - w_0^\nmh}{\myhalf(\dt^n+\dt^{n-1})} 
+ w_0^n \left(\frac{\partial w_0}{\partial r}\right)^n.
\end{equation}
where $w_0^n$ and $(\partial w_0 / \partial r)^{n}$ are defined through paper III equation (53),
\begin{equation}
w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}\enskip,\label{eq:w0nstar}
\end{equation}
and paper III equation (54),
\begin{equation}
\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] \enskip.\label{eq:dw0drnstar}
\end{equation}

@@

The procedure to construct $\uadvtwo$ is described in detail in Appendix B of paper III
and is analogous to the procedure used in step 2, but with updated values
for $w_0$ and $\pizero.$  We note that $\uadvtwo$ satisfies the discrete versions of 
equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}).   The discretization
with the volume discrepancy correction is
\begin{eqnarray}
\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) &=&
\beta_0^{\nph,\star}\left\{\left(S^{\nph} - \overline{S^{\nph}}\right)
 + \frac{f}{\gammabar^{\nph,\star} p_0^{\nph,\star}}\cdot \right. \nonumber \\
&& \left.\left[\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right]\right\} \enskip ,
\end{eqnarray}
\begin{equation}
\beta_0^{\nph,\star} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2} \enskip .
\end{equation}

@@

\item Update $p_0$ and compute $\psi$.
\begin{description}
\item For plane-parallel,\\
{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph}, \etarho^{n+\myhalf}]$ 
$\rightarrow [p_0^{n+1},\psi^{\nph}]$.
\item For spherical,\\
{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{n+1,\star},\overline{S^{\nph,\star}}] \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
\end{description}

@@

The velocity update happens analogously to paper II, using $S^{n+1}$ from step 10.
We update the velocity field $\Ubt^n$ to $\Ubt^{n+1,\star}$ by discretizing 
equation (\ref{eq:utildeupd}) as
\begin{eqnarray}
\Ubt^{n+1,\star} &=& \Ubt^n - \dt \;
 \left[\left(\uadvtwo+ w_0^{\nph} \er\right) \cdotb \nablab \Ubt \right]^\nph
      - \dt \; \left(\uadvtwo \cdotb \er\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \er \nonumber \\
   &&   + \dt \left[ - \frac{1}{\rho^\nph} {\mathbf{G}} \pi^\nmh
        + \frac{1}{\rho_0} {\mathbf{G}} \pizerotwo
        - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g \er \right] \enskip ,
\end{eqnarray}
where $\rho^\nph = ( \rho^n + \rho^{n+1} )/2$
and $\mathbf{G}$ approximates a cell-centered gradient from nodal data. 
The construction of $[(\uadvtwo+ w_0^{\nph} \er \cdotb \nablab) \Ubt ]^\nph$
is described in paper II, Appendix A, with $\Vb = \uadvtwo+w_0^{\nph} \er$ and~$s$ set to each
component of $\Ubt^n$ individually.

Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}) 
including the volume discrepancy correction,
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph} \Ubt^{n+1} \right)  = \beta_0^{\nph}\left\{  \left(S^{n+1} - \overline{S^{n+1}} \right)
+ \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]\right\}
\enskip,
\end{equation}
by defining ${\bf V} = \Ubt^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
and solving
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph}\left\{ \left(S^{n+1}-\overline{S^{n+1}} \right)
- \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]\right\} \enskip, \label{Linear System for Multigrid}
\end{equation}
for nodal values of $\phi$, where
$\beta_0^\nph = ( \beta_0^n + \beta_0^{n+1} )/2$ and
$L_\beta^\rho$ is the standard bilinear
finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
In this step, $D$ is a discrete second-order operator that approximates the 
divergence at nodes from cell-centered data and satisfies
$\mathbf{G} = -D^T.$ 
(See (cite almgrenBellSzymczak:1996) for a detailed discussion of this
approximate projection; see (cite almgren:bell:crutchfield) for a discussion
of this particular form of the projection operand.)  
We solve the linear system of equation (\ref{Linear System for Multigrid})
using multigrid V-cycles with Gauss-Seidel relaxation.

>>>>

\section{Changes Between Paper 3 and Paper 4}
\begin{enumerate}
\item {\tt fill\_3d} and {\tt average} algorithms.
\item We update temperature after each {\bf React State} as well as the advection step.
\item We have created a {\tt burning\_cutoff\_density}, where the burning does
not happen below this density.  It is presently set to {\tt base\_cutoff\_density}.
\item Use corner coupling in advection.
\item We use {\tt use\_tfromp} to update temperature using $T=T(\rho,X_k,p_0)$ rather 
than $T=T(\rho,h,X_k)$.
\end{enumerate}
Changes which modify enthalpy update.  These are separate because in this paper,
enthalpy is only used as a diagnostic since we use {\tt use\_tfromp}.  If we don't
mention these items in paper IV, they should me noted in paper 4.5.
\begin{enumerate}
\item We have changed the discretization of $\Ub\cdot\nabla p_0$ in the enthalpy
update to $\nabla\cdot(\Ub p_0) - p_0\nabla\cdot\Ub$.
\item In paper III we discretized the enthalpy evolution equation in
full, non-perturbation form, $\rho h$.  Since then we have discovered that 
discretizing the enthalpy evolution in perturbational form, $(\rho h)'$,
leads to better numerical properties.  We use {\tt enthalpy\_pred\_type = 1}.
This is more like paper II.
\item We have turned off the evolution of $h$ above the atmosphere and instead
compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff = TRUE}.
\end{enumerate}
\section{Changes Between Papers 4 and Papers 4.5}
\begin{enumerate}
\item We split the update of the density and enthalpy in steps 4 and 8.
Now, density is updated first, allowing us to compute time-centered values of 
$\psi$ that are then used in the enthalpy update.
\item We split {\bf Advect Base} into separate {\bf Advect Base Density},
{\bf Advect Base Pressure}, and {\bf Advect Base Enthalpy}.  The calling sequence
is in the main flow chart.
\item We replaced {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
plane-parallel and $\rho_0 = \rho_0 + {\rm{\bf Avg}}(\rho - \rho_0)$ in spherical,
eliminating the need for $\nabla\cdot\etarho$.  {\bf Correct Base} is now used
analogously for setting $(\rho h)_0$ after {\bf React State}.
\item After {\bf React State}, we set $(\rho h)_0 = {\rm{\bf Avg}}(\rho h)$ in 
plane-parallel and $(\rho h)_0 = (\rho h)_0 + {\rm{\bf Avg}}[(\rho h) - (\rho h)_0]$ 
in spherical.  This means that {\bf React Base} no longer exists, and the intermediate 
states, $(\rho h)_0^{(1)}$ and $(\rho h)_0^{(2)}$, no longer have to exist.  
{\bf Advect Base Enthalpy} will advance $(\rho h)_0^n \rightarrow (\rho h)_0^{n+1}$ 
directly.  This also means we don't have to deal with $\eta_{\rho h}$ or a 
{\bf Diffuse Base}.
\item In {\bf Step 4} and {\bf Step 8}, we moved {\bf Correct Base} so it
occurs after the full state density update and  before {\bf Advect Base Pressure}.  
Since the full enthalpy advance ``feels'' the effects of the time-centered $\etarho$, 
we should rely on a $\rho_0^{n+1}$ that also feels the effects.
\item {\bf AMR!!!}
\end{enumerate}
Changes specific to spherical:
\begin{enumerate}
\item Discuss {\bf Advect Base} routines and {\bf Correct Base} routine.
\item $\etarho$ and $\psi$ equations.
\end{enumerate}
\section{Changes Between Papers 4.5 and Paper 5}
\begin{enumerate}
\item Use PPM limiters in advection.
\end{enumerate}
\section{Changes Between Paper 5 and XRB Paper}
\begin{enumerate}
\item We have added thermal diffusion, controlled by {\tt use\_thermal\_diffusion},
{\tt temp\_diffusion\_formulation}, and {\tt thermal\_diffusion\_type}.
\item We added the volume discrepancy term to the velocity constraint equation,
controlled by the input parameter, {\tt dpdt\_factor}.
\end{enumerate}

@@

We integrate
\begin{equation}
\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \psi
\enskip \label{eq:divw0} ,
\end{equation}
with the following discretization:
\begin{equation}
\frac{\partial w_0^{\outp}}{\partial r} =  \Sbar^{\inp}
- \frac{1}{\gammabar^{\inp} p_0^{\inp}}\psi^{\inp}
\enskip ,
\end{equation}
and $w_0=0$ at $r=0$.\\

The discretization with the volume discrepancy correction require an additional argument,
$\overline{p^{\inp}}$, and the discretization becomes:
\begin{equation}
\frac{\partial w_0^{\nph,\outp}}{\partial r} = 
\overline{S^{\nph,\inp}} - \frac{1}{\gammabar^{\inp} p_0^{\inp}}
\left[ \psi^{\inp} + f \left(\frac{p_0^{\inp} - \overline{p^{\inp}}}{\Delta t^n}\right)\right]
\enskip .
\end{equation}

@@

If this is the first step of the calculation ($n=0$), we set
\begin{equation}
S^{\myhalf,\star} = \frac{S^0 + S^1}{2} \enskip ,
\end{equation}
where $S^1$ is found through the iterative process that initializes the calculation.
Otherwise, following the method used in our small-scale low Mach number algorithm
(BDRWZ), we extrapolate to the half-time using the source
terms at the previous and current time levels
\begin{equation}
S^{\nph,\star} = S^n + \frac{\Delta t^n}{2} \frac{S^n - S^{n-1}}{\Delta t^{n-1}} \enskip .
\end{equation}
\item Compute
\begin{equation}
\overline{S^{\nph,\star}} = {\mathrm{\bf Avg}} \left(S^{\nph,\star}\right) \enskip .
\end{equation}
\item Compute $w_0^{n+\myhalf,\star}$:
\begin{description}
\item For plane-parallel,\\
{\bf Compute $w_0$ Planar}$[\overline{S^{\nph,\star}},\gammabar^n,p_0^n,\psi^{n-\myhalf}] 
\rightarrow [w_0^{n+\myhalf,\star}]$.

If we are using the volume discrepancy term, compute 
$\overline{p^{\inp}} = \overline{p(\rho,h,X_k)^n}$.
\item For spherical,\\
{\bf Compute $w_0$ Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,\rhozero^n,p_0^n,
\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.
\end{description}

@@

The procedure to construct $\uadvone$ is described in detail in Appendix B of paper III.
We note that  $\uadvone$ satisfies the discrete versions of paper III equation (21),
\begin{equation}
\int_{\Omega_H} \Ubt \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
\end{equation}
and paper III equation (39),
\begin{equation}
\nablab \cdotb (\beta_0 \Ubt )  = \beta_0 \left(S - \Sbar \right )\enskip ,
\label{eq:tildeconstraint}
\end{equation}
with the discretization
\begin{equation}
\nablab \cdotb \left(\beta_0^n \uadvone\right) = 
\beta_0^n \left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right) \enskip .
\end{equation}
The discretization with the volume discrepancy correction is
\begin{equation}
\nablab \cdotb \left(\beta_0^n \uadvone\right) = 
\beta_0^n \left\{ \left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right)
+ \frac{f}{\gammabar^n p_0^n}
\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right]\right\} \enskip .
\end{equation}

@@

\item Define:
\begin{equation}
\gammabar^{\nph} = \frac{\gammabar^{n} + \gammabar^{n+1,\star}}{2}\enskip , ~~~
\rhozero^{\nph} = \frac{\rhozero^{n} + \rhozero^{n+1,\star}}{2}\enskip , ~~~
p_0^{\nph} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}\enskip .
\end{equation}
\item Compute $w_0^{n+\myhalf}$:
\begin{description}
\item For plane-parallel,\\
{\bf Compute $w_0$ Planar}$[\overline{S^{\nph}},
\gammabar^{n+\myhalf},p_0^{n+\myhalf},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf}]$.

If we are using the volume discrepancy term, compute
\begin{equation}
\overline{p^{\inp}} = \frac{\overline{p(\rho,h,X_k)^n} + \overline{p(\rho,h,X_k)^{n+1,\star}}}{2}.
\end{equation}
\item For spherical,\\
{\bf Compute $w_0$ Spherical}$[\overline{S^{\nph}},\gammabar^{n+\myhalf},
\rhozero^{n+\myhalf},p_0^{n+\myhalf},\etarho^{n+\myhalf}] \rightarrow [w_0^{n+\myhalf}]$.
\end{description}

\item Using equation (\ref{eq:pizero}), define the scaled pressure gradient using the
following discretization:
\begin{equation}
-\left ( \frac{1}{\rho_0^n} \frac{\partial \pizerotwo}{\partial r} \right ) = 
\frac{w_0^{\nph} - w_0^\nmh}{\myhalf(\dt^n+\dt^{n-1})} 
+ w_0^n \left(\frac{\partial w_0}{\partial r}\right)^n.
\end{equation}
where $w_0^n$ and $(\partial w_0 / \partial r)^{n}$ are defined through paper III equation (53),
\begin{equation}
w_0^n = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph}}{\dt^n+\dt^{n-1}}\enskip,\label{eq:w0nstar}
\end{equation}
and paper III equation (54),
\begin{equation}
\left(\frac{\partial w_0}{\partial r}\right)^{n} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph} \right ] \enskip.\label{eq:dw0drnstar}
\end{equation}

@@

The procedure to construct $\uadvtwo$ is described in detail in Appendix B of paper III
and is analogous to the procedure used in step 2, but with updated values
for $w_0$ and $\pizero.$  We note that $\uadvtwo$ satisfies the discrete versions of 
equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}), with the discretization
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph} \uadvtwo\right) =
\beta_0^{\nph}\left(S^{\nph} - \overline{S^{\nph}}\right) \enskip ,
\end{equation}
\begin{equation}
\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2} \enskip .
\end{equation}
The discretization with the volume discrepancy correction is
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph} \uadvtwo\right) =
\beta_0^{\nph}\left\{\left(S^{\nph} - \overline{S^{\nph}}\right)
 + \frac{f}{\gammabar^{\nph} p_0^{\nph}}\cdot 
\left[\frac{p(\rho,h,X_k)^{\nph} - \overline{p(\rho,h,X_k)^{\nph}}}{\Delta t^n}\right]\right\} \enskip ,
\end{equation}

@@

\item Update $p_0$ and compute $\psi$.
\begin{description}
\item For plane-parallel,\\
{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph}, \etarho^{n+\myhalf}]$ 
$\rightarrow [p_0^{n+1},\psi^{\nph}]$.
\item For spherical,\\
{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{n+1,\star},\overline{S^{\nph}}] \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
\end{description}

@@

The velocity update happens analogously to paper II, using $S^{n+1}$ from step 10.
We update the velocity field $\Ubt^n$ to $\Ubt^{n+1,\star}$ by discretizing 
equation (\ref{eq:utildeupd}) as
\begin{eqnarray}
\Ubt^{n+1,\star} &=& \Ubt^n - \dt \;
 \left[\left(\uadvtwo+ w_0^{\nph} \er\right) \cdotb \nablab \Ubt \right]^\nph
      - \dt \; \left(\uadvtwo \cdotb \er\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \er \nonumber \\
   &&   + \dt \left[ - \frac{1}{\rho^\nph} {\mathbf{G}} \pi^\nmh
        + \frac{1}{\rho_0} {\mathbf{G}} \pizerotwo
        - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g \er \right] \enskip ,
\end{eqnarray}
where $\rho^{\nph}$ is defined as
\begin{equation}
\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}\enskip ,
\end{equation}
and $\mathbf{G}$ approximates a cell-centered gradient from nodal data. 
The construction of $[(\uadvtwo+ w_0^{\nph} \er \cdotb \nablab) \Ubt ]^\nph$
is described in paper II, Appendix A, with $\Vb = \uadvtwo+w_0^{\nph} \er$ and~$s$ set to each
component of $\Ubt^n$ individually.

Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}):
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph} \Ubt^{n+1} \right) 
= \beta_0^{\nph} \left(S^{n+1} - \overline{S^{n+1}} \right) \enskip,
\end{equation}
where $\beta_0^{n+\myhalf}$ is now defined as
\begin{equation}
\beta_0^{n+\myhalf} = \frac{\beta_0^n + \beta_0^{n+1}}{2}\enskip .
\end{equation}
The discretization with the volume discrepancy term is:
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph} \Ubt^{n+1} \right)  = \beta_0^{\nph}\left\{  \left(S^{n+1} - \overline{S^{n+1}} \right)
+ \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]\right\}
\enskip .
\end{equation}

To solve this equation, we define
${\bf V} = \Ubt^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
and solve
\begin{equation}
 L_\beta^\rho \phi = D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} 
\left(S^{n+1}-\overline{S^{n+1}} \right) \enskip, \label{Linear System for Multigrid}
\end{equation}
for nodal values of $\phi$, where $L_\beta^\rho$ is the standard bilinear
finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
The discretization with the volume discrepancy term is:
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph}\left\{ \left(S^{n+1}-\overline{S^{n+1}} \right)
- \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]\right\} \enskip.
\end{equation}
In this step, $D$ is a discrete second-order operator that approximates the 
divergence at nodes from cell-centered data and satisfies
$\mathbf{G} = -D^T.$ 
(See (cite almgrenBellSzymczak:1996) for a detailed discussion of this
approximate projection; see (cite almgren:bell:crutchfield) for a discussion
of this particular form of the projection operand.)  
We solve the linear system of equation (\ref{Linear System for Multigrid})
using multigrid V-cycles with Gauss-Seidel relaxation.
